---
title: "P8105 Homework 4"
date: "Last Updated by K. Frierson 2024-10-26"
output: 
  html_document:
    toc: true
    toc_float: true
---
# Introduction

The fourth homework assignment for P8105: Data Science I requires the build of a website (where you are viewing this right now) and the implementation of a `plotly` dashboard. I will be using NYC restraurant inspection data from the `p8105.datasets` package.

# Call Libraries

Load packages needed for entire assignment:
```{r libraries, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(knitr)
library(leaflet)
library(sf)
```

# Data Import & Cleaning

Bring in data: 
```{r data}
data(rest_inspec)
```

There is a lot of data here which would hinder easy visualization. I will drop rows with missing and keep only variables I am interested in. Also I will create a `violation_severity` variable  for the corresponding inspection score based on definitions given on the data science course [website](https://p8105.com/dataset_restaurant_inspections.html): 
```{r}
inspect = 
  rest_inspec |>
  mutate(
    violation_severity = case_when(
      score >= 7 ~ "Public Health Hazard",
      score >= 5 ~ "Critical Violation",
      score >= 2 ~ "General Violation",
      TRUE ~ "No Violation"
    )
  ) |> 
  select(boro, camis, critical_flag, violation_severity,
         cuisine_description, dba, inspection_date, score, street,
         violation_code, violation_description, zipcode, grade, grade_date) |>   
  drop_na() 
```


# EDA: Exploratory Data Analysis

How many of each type of violation?
```{r}
violation_summary =
  inspect |>
  group_by(violation_severity) |>  
  summarize(Count = n())  |> 
  filter(violation_severity %in% c("Public Health Hazard", 
                                   "Critical Violation", "General Violation"))

kable(violation_summary)

```
