---
title: "P8105 Homework 4"
date: "Last Updated by K. Frierson 2024-10-26"
output: 
  html_document:
    toc: true
    toc_float: true
---
# Introduction

The fourth homework assignment for P8105: Data Science I requires the build of a website (where you are viewing this right now) and the implementation of a `plotly` dashboard. I will be using NYC restraurant inspection data from the `p8105.datasets` package.

# Call Libraries

Load packages needed for entire assignment:
```{r libraries, message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(knitr)
library(leaflet)
library(sf)
library(viridis)
library(plotly)
```

# Data Import & Cleaning

Bring in data: 
```{r data}
data(rest_inspec)
```

There is a lot of data here which would hinder easy visualization. I will drop rows with missing and keep only variables I am interested in. Also I will create a `violation_severity` variable  for the corresponding inspection score based on definitions given on the data science course [website](https://p8105.com/dataset_restaurant_inspections.html): 
```{r}
inspect = 
  rest_inspec |>
  mutate(
    violation_severity = case_when(
      score >= 7 ~ "Public Health Hazard",
      score >= 5 ~ "Critical Violation",
      score >= 2 ~ "General Violation",
      TRUE ~ "No Violation"
    )
  ) |> 
  select(boro, camis, critical_flag, violation_severity,
         cuisine_description, dba, inspection_date, score, street,
         violation_code, violation_description, zipcode, grade, grade_date) |>
  drop_na() 
```


# Descriptive Analysis

How many of each type of violation?
```{r}
violation_summary =
  inspect |>
  group_by(violation_severity) |>  
  summarize(Count = n())  |> 
  filter(violation_severity %in% c("Public Health Hazard", 
                                   "Critical Violation", "General Violation"))

kable(violation_summary)

```

# Plots

Lets show distribution of scores, use `vidiris` colors for maximum accessibility: 
```{r, message=FALSE, warning=FALSE}
histogram =
  inspect |> 
  filter(violation_severity != "No Violation") |> 
   mutate(violation_severity = fct_relevel(violation_severity, "General Violation",
                                           "Critical", "Public Health Hazard")) |> 
  ggplot(aes(x=score, fill=violation_severity))+
  geom_histogram(alpha=0.9, binwidth = 1.9) +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  xlim(0, 70)+
  labs(x= "Health Inspection Score", y = "Count", fill = "Violation Severity")
    

histogram
```

Lets compare previous health grade to inspection score: 
```{r}

previously_graded =
  inspect |>
  mutate(
    grade_date = as.Date(grade_date),
    inspection_date = as.Date(inspection_date)
  )

previoulsy_graded = 
  inspect |> 
  filter(grade_date < inspection_date)

grade_v_score =
  previously_graded |>
  filter(!grade %in% c("Not Yet Graded", "P", "Z")) |> 
  ggplot(aes(x= grade, y=score)) +
  geom_boxplot() +
  labs(x= "Previous Health Grade", y = "Latest Inspection Score")+
  theme_classic()

grade_v_score
  
```

lets map average score by zip code: 
```{r}

nyc_zip_codes = st_read("homework4_files/Modified Zip Code Tabulation Areas (MODZCTA)/geo_export_e426608c-b383-45c5-8f28-cc94eb8cc12f.shp") |>
  rename(zipcode = modzcta) |> 
    mutate(zipcode = as.integer(zipcode))|>
  st_transform(crs = 4326)

avg_score_zip = 
  inspect |> 
  group_by(zipcode) |> 
  summarize(avg_score = mean(score, na.rm = TRUE))

merged_geo_data = 
  nyc_zip_codes |>
  left_join(avg_score_zip, by = "zipcode")

pal = 
  colorNumeric(palette = "YlOrRd", domain = merged_geo_data$avg_score, 
               na.color = "transparent")

map = 
merged_geo_data |> 
leaflet()|> 
  addTiles() |> 
  addPolygons(
    fillColor = ~pal(avg_score),      
    fillOpacity = 0.7,                    
    color = "white",                       
    weight = 1,                            
    highlightOptions = highlightOptions(
      weight = 5,
      color = "blue",
      fillOpacity = 0.9,
      bringToFront = TRUE
    ),
    label = ~paste("Zip Code:", zipcode, "Average Score:", round(avg_score, 2)),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", "font-size" = "12px"), 
      textsize = "15px",
      direction = "auto"
    )
  ) |> 
  addLegend(pal = pal, values = ~avg_score, opacity = 0.7, 
            title = "Average Score", position = "bottomright")

map
```
